Simulations of Ballot Polling Risk-Limiting Audits

### 2021-11-02 MultipleRoundSims/draft.tex
definitions

\begin{document}
\lstset{language=Python}
%
\title{Simulations of Ballot Polling Risk-Limiting Audits}
%add \thanks{} within the title brackets if want to refer to supporting organization ^^
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\begin{comment}
\author{Oliver Broadrick\inst{1} \and
Sarah Morin\inst{1} \and
Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
Filip Zag{\'o}rski\inst{3} }
%
\authorrunning{Broadrick, Morin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Computer Science, The George Washington University 
\email{obroadrick@gwmail.gwu.edu}\\
\and
Department of Computer Science, Stanford University 
\email{grantmcc@stanford.edu}\\
\and Wroclaw University of Science and Technology\\
\email{filip.zagorski@gmail.com}}
%
\end{comment}
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling risk limiting audits (RLAs) \Minerva, Selection-Ordered \BRAVO and End-of-Round \BRAVO. We also
    present details on the R2B2 open source software library and the related simulation software.

    \BRAVO is the most commonly used ballot-polling RLA. \Minerva was recently proposed, and requires fewer first-round ballots, on average, than both Selection-Ordered \BRAVO and End-of-Round \BRAVO when the first-round stopping probability is large.

    An open question, however, is how \Minerva compares to Selection-Ordered \BRAVO and End-of-Round \BRAVO over multiple rounds.  In this paper, we present results from simulations of multiple round audits with first-round stopping probabilities of $90\%$, a common choice among election officials. Because the size of rounds in \Minerva needs to be predetermined (and independent of the audit draws), we use two pre-determined round sequences for \Minerva: (a) all rounds are of the same size (which might be a practical choice for election officials if they had planned to have resources for the first round size) and (b) each round is $1.5$ times the previous one (which is the preset value for \Minerva as integrated into election audit software Arlo).

@@

\keywords{Risk-limiting audit  \and Ballot polling audit}
\end{abstract}
%
%
%
\section{Introduction}
\label{sec:intro}
\input{sections/01-intro}

>>>>

\begin{document}
\lstset{language=Python}
%
\title{Simulations of Ballot Polling Risk-Limiting Audits}
%add \thanks{} within the title brackets if want to refer to supporting organization ^^
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\begin{comment}
\author{Oliver Broadrick\inst{1} \and
Sarah Morin\inst{1} \and
Grant McClearn\inst{2} \and Neal McBurnett \and Poorvi L. Vora\inst{1} \and
Filip Zag{\'o}rski\inst{3} }
%
\authorrunning{Broadrick, Morin et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Department of Computer Science, The George Washington University 
\email{obroadrick@gwmail.gwu.edu}\\
\and
Department of Computer Science, Stanford University 
\email{grantmcc@stanford.edu}\\
\and Wroclaw University of Science and Technology\\
\email{filip.zagorski@gmail.com}}
%
\end{comment}
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    In this paper we present simulation results comparing the risk, stopping probability and number of ballots required over multiple rounds of ballot-polling risk limiting audits (RLAs) \Minerva, Selection-Ordered (SO) \BRAVO and End-of-Round (EoR) \BRAVO. We also
    present details on the R2B2 open source software library and the related simulation software.

    \BRAVO is the most commonly used ballot-polling RLA and requires the smallest expected number of ballots when ballots are drawn one at a time and the (true) underlying election is as announced. In real audits, election officials draw multiple ballots at a time, and \BRAVO rules may be implemented as SO \BRAVO or EoR \BRAVO, with neither being optimal for large first round sizes. \Minerva was recently proposed for use when the first-round stopping probability is large. It requires fewer first-round ballots, on average, than both SO \BRAVO and EoR \BRAVO for first round stopping probabilities of $0.9$ when the (true) underlying election is as announced.

    An open question, however, is how \Minerva compares to SO \BRAVO and EoR \BRAVO over multiple rounds, and for lower stopping probabilities.  In this paper, we present results from simulations of multiple round audits. We examine both: first-round stopping probabilities of $90\%$, a common choice among election officials, and $25\%$, which would be more favorable to \BRAVO. Because the size of rounds in \Minerva needs to be predetermined (and independent of the audit draws), we use two pre-determined round sequences for \Minerva: (a) all rounds are of the same size (which might be a practical choice for election officials if they had planned to have resources for the first round size) and (b) each round is $1.5$ times the previous one (which is the preset value for \Minerva as integrated into election audit software Arlo).

@@

\keywords{risk-limiting audit (RLA)  \and ballot polling audit \and evidence-based elections \and statistical election audit}
\end{abstract}
%
%
%
\section{Introduction}
\label{sec:intro}
\input{sections/01-intro}

### 2021-11-02 MultipleRoundSims/sections/01-intro.tex
definitions

%Intro
The literature contains numerous descriptions of vulnerabilities in deployed voting systems, and it is not possible to be certain that any system, however well-designed, will perform as expected in all instances. For this reason, 
{\em evidence-based elections} aim to produce trustworthy and compelling evidence of the correctness of election outcomes, enabling the detection of problems with high probability. One way to implement an evidence-based election is to use a well-curated voter-verified paper trail, compliance audits, and a rigorous tabulation audit of the election outcome, known as a risk-limiting audit (RLA). This paper provides insight into the main approach to ballot polling RLAs, the \BRAVO audit \cite{bravo}, and the newer \Minerva \cite{usenix_minerva} ballot polling RLA, through the presentation of simulation results. While some properties of the two audits may be theoretically derived, for other properties theoretical results are not available. This paper examines the number of ballots drawn over multiple rounds of both audits, and also examines the related probabilities of stopping if the election is as announced, and the maximum risk of the audit.

\subsection{Background}
This paper focuses on ballot-polling RLAs, which require a large number of ballots but do not rely on any special features of the election technology. In the general ballot-polling risk limiting audit (RLA), a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}. A statistical measure is then computed to determine whether there is sufficient evidence to declare the election outcome correct within the pre-determined risk limit. Because the decision is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit. The special case when round size is one---that is, stopping decisions are made after each ballot draw---is a {\em ballot-by-ballot (B2)} audit.

@@

\Minerva was designed for R2 audits and applies its stopping rule once for each round. Thus it does not require the tracking of ballots that SO \BRAVO does. It has been proven that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds. Simulations have also been presented to understand better its first round properties; these show that \Minerva also draws fewer ballots than SO \BRAVO for large first rounds. There are no results, however, either theoretical or based on simulations, regarding the expected number of ballots drawn over multiple rounds in a \Minerva audit. Further, there is no literature comparing the number of ballots drawn by \Minerva and EoR or SO \BRAVO over multiple rounds.

@@

\subsection{Organization}

>>>>

%Intro
The literature contains numerous descriptions of vulnerabilities in deployed voting systems, and it is not possible to be certain that any system, however well-designed, will perform as expected in all instances. For this reason, 
{\em evidence-based elections} \cite{evidence-based} aim to produce trustworthy and compelling evidence of the correctness of election outcomes, enabling the detection of problems with high probability. One way to implement an evidence-based election is to use a well-curated voter-verified paper trail, compliance audits, and a rigorous tabulation audit of the election outcome, known as a risk-limiting audit (RLA) \cite{RLA}. An RLA is an audit which guarantees that the probability of concluding that an election outcome is correct, given that it is not, is below a pre-determined value known as the risk limit of the audit, independent of the true, unknown vote distribution of the underlying election. Over a dozen states have seriously explored the use of RLAs---some have pilot programs, some allow RLAs to satisfy a general audit requirement and some have RLAs in statute.

This paper provides insight into the main approach to ballot polling RLAs, the \BRAVO audit \cite{bravo}, and the newer \Minerva \cite{usenix_minerva} ballot polling RLA, through the presentation of simulation results. While some properties of the two audits may be theoretically derived, for other properties theoretical results are not available. This paper examines the number of ballots drawn over multiple rounds of both audits, the related probabilities of stopping if the election is as announced, and the maximum risk of the audit.

\subsection{Background}
Of the multiple types of RLAs, this paper focuses on ballot-polling RLAs, which require a large number of ballots but do not rely on any special features of the election technology. In the general ballot-polling RLA, a number of ballots are drawn and tallied in what is termed a {\em round} of ballots \cite{usenix_minerva}. A statistical measure is then computed to determine whether there is sufficient evidence to declare the election outcome correct within the pre-determined risk limit. Because the decision is made after drawing a round of ballots, the audit is termed a {\em round-by-round (R2)} audit. The special case when round size is one---that is, stopping decisions are made after each ballot draw---is a {\em ballot-by-ballot (B2)} audit.

@@

\Minerva was designed for R2 audits and applies its stopping rule once for each round. Thus it does not require the tracking of ballots that SO \BRAVO does. Zag{\'o}rski {\em et al.} \cite{usenix_minerva} prove that \Minerva is a risk-limiting audit and requires fewer ballots to be sampled than EoR \BRAVO when an audit is performed in rounds and the round schedule is pre-determined (before any ballots are drawn). They also present first-round simulations which show that \Minerva draws fewer ballots than SO \BRAVO in the first round for large first rounds when the (true) underlying election is as announced.

There are no results, either theoretical or based on simulations, regarding the number of ballots drawn over multiple rounds in a \Minerva audit with a pre-determined schedule. Because \BRAVO does not need to work on a pre-determined round schedule, it can optimize the size of the next round based on the sample drawn so far. Thus an open question is whether the constraint of a predetermined round schedule limits the efficacy of \Minerva in future rounds, and there is no literature comparing the number of ballots drawn by \Minerva and SO \BRAVO over multiple rounds. Note that the Average Sample Number (ASN) computations for \BRAVO \cite{bravo} apply only for B2 audits and are especially misleading as estimates of the number of ballots drawn over multiple rounds when first round sizes are large.

@@

\subsection{Organization} Section \ref{sec:related} describes related work and \ref{sec:software} our open source software. The experiments we performed are described in section \ref{sec:expts} and sections \ref{sec:results1} and \ref{sec:results2} present our results. Section \ref{sec:conc} has our conclusions.

### 2021-11-02 MultipleRoundSims/sections/02-related.tex
definitions

%Related Work
The \BRAVO audit \cite{bravo} is a well-known ballot polling audit which has been used in numerous pilot and real audits. When used to audit a two-candidate election, it is an instance of Wald's sequential probability ratio test (SPRT) \cite{wald}, and inherits the SPRT property of being the most efficient test (requiring the smallest expected number of ballots) if the election is as announced. The model for \BRAVO and the SPRT is, however, that of a sequential audit: a sample of size one is drawn, and a decision of whether to stop the audit or not is taken. Real election audits invest in drawing a large number of ballots before making the decision. It is possible to apply \BRAVO to the sequence of ballots if the sequential order is retained. This is not, however, the most efficient possible use of the drawn sample because information in consequent ballots is ignored when applying \BRAVO to a set of ballots drawn earlier.

>>>>

%Related Work
The \BRAVO audit \cite{bravo} is a well-known ballot polling audit which has been used in numerous pilot and real audits. When used to audit a two-candidate election, it is an instance of Wald's sequential probability ratio test (SPRT) \cite{wald}, and inherits the SPRT property of being the most efficient test (requiring the smallest expected number of ballots) if the election is as announced. The model for \BRAVO and the SPRT is, however, that of a sequential audit: a sample of size one is drawn, and a decision of whether to stop the audit or not is taken. Real election audits invest in drawing a large number of ballots before making the decision. It is possible to apply \BRAVO to the sequence of ballots if the sequential order is retained. This is not, however, the most efficient possible use of the drawn sample because information in consequent ballots is ignored when applying \BRAVO to ballots that were drawn earlier in the sample.

### 2021-11-02 MultipleRoundSims/sections/03-software.tex
definitions

The \code{Audit} contains information related to the audit parameters for a single contest, 
such as the risk limit, sampling method, and \code{Contest} to audit. It is important
to note the \code{Audit} is an Abstract Base Class upon which specific RLAs are built.
It only contains the parameters and attributes common to the RLAs of this paper and provides a set
of methods that can be called by any audit implementation. The functionality of
\code{Audit} can be divided into two basic groups: \textit{interactive}
and \textit{bulk}.\\
\\
The interactive implementation allows users to execute an audit step-by-step as it
might progress during a live election audit through the following:

@@

R2B2 also provides a significant amount of functionality `out-of-the-box' for educational
or exploratory use. For those who wish learn about RLAs without having to write any
code themselves, R2B2 provides a command line tool for both interactive auditing and
generating audit results and statistics for larger data sets.

\subsection{Simulation Software}
As described above R2B2 has implementations of several ballot polling risk-limiting audits as well as a simulator, 
all written in Python.
For each of these audits, the software can compute the stopping condition for a given sample and estimates
of the next round size to achieve a desired stopping probability. 
For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator, [need to check].
given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be those announced).

>>>>

The \code{Audit} contains information related to the audit parameters for a single contest, 
such as the risk limit, sampling method, and \code{Contest} to audit. It is important
to note the \code{Audit} is an Abstract Base Class upon which specific RLAs are built.
It only contains the parameters and attributes common to the RLAs of this paper and provides a set
of methods that can be called by any audit implementation. The functionality of
\code{Audit} can be divided into two basic groups: \textit{interactive}
and \textit{bulk}.

The interactive implementation allows users to execute an audit step-by-step as it
might progress during a live election audit through the following:

@@

R2B2 also provides a significant amount of functionality `out-of-the-box' for educational
or exploratory use. For those who wish to learn about RLAs without having to write any
code themselves, R2B2 provides a command line tool for both interactive auditing and
generating audit results and statistics for larger data sets.

\subsection{Simulation Software}
As described above R2B2 has implementations of several ballot polling risk-limiting audits as well as a simulator, 
all written in Python.
For each of these audits, the software can compute the stopping condition for a given sample and estimates
of the next round size to achieve a desired stopping probability. 
For a given audit and random seed, the simulator draws random samples, with replacement, using a pseudorandom number generator.
given the number of votes for each candidate, and the number of invalid votes, in the underlying election (these need not be chosen to be those announced).

### 2021-11-02 MultipleRoundSims/sections/04-expts.tex
definitions

All of the audits discussed in this paper are modeled as binary hypothesis tests.

Under the alternate hypothesis, $H_a$, the announced outcome is correct. 
That is, the true underlying ballot distribution is given by the announced ballot tallies.

Under the null hypothesis, $H_0$, the true outcome is a tie 
(or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the winner).

@@

It is useful to discuss the probability of an audit stopping in
the $j^{th}$ round, given that the underlying election is as announced.
\begin{definition}[Stopping Probability]
The stopping probability $S$ of an audit $\mathcal{A}$ in round $j$ is 
$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
\end{definition}
The notion of stopping probability can be useful for selecting round sizes.

We performed simulations to study $S_j$ and obtain estimates of the expected number of ballots drawn for the various audits. We know that \BRAVO uses the smallest expected number of ballots when drawn ballot-by-ballot. However, this may not be the case when \BRAVO is used as an R2 audit for a large first round size. We also studied the risk values. While all the audits we studied have been proven to be RLAs, it is useful to observe how close to the risk limit each audit gets over a small, finite number of rounds as would be the case in a real election.

>>>>

All of the audits discussed in this paper are modeled as binary hypothesis tests. Under the alternate hypothesis, $H_a$, the announced outcome is correct. That is, the true underlying ballot distribution is given by the announced ballot tallies. Under the null hypothesis, $H_0$, the true outcome is a tie (or the announced winner lost by one vote, and the number of ballots is large enough that the probability of drawing a ballot for the winner is that of drawing one for the winner).

@@

We present measures of stopping probability in the $j^{th}$ round of the audit, given that the underlying election is as announced.
\begin{definition}[Stopping Probability]
The stopping probability $S_j$ of an audit $\mathcal{A}$ in round $j$ is 
$$S_j(\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\land \mathcal{A}(X) \neq Correct ~previously \mid H_a]$$
\end{definition}
Experimentally, using our simulations, $S_j$ would be estimated by the fraction of audits that stop in round $j$.

Note that $\sum _j S_j(\mathcal{A})=1$. We can also consider the cumulative stopping probability: 
\begin{definition}[Cumulative Stopping Probability]
The cumulative stopping probability $C_j$ of an audit $\mathcal{A}$ in round $j$ is 
$$C_j(\mathcal{A})= \sum_{i=1}^j S_j$$
\end{definition}
Experimentally, using our simulations, $C_j$ would be estimated by the fraction of audits that stop in or before round $j$.

Finally, we are also interested in the probability that an audit will stop in round $j$ given that it did not stop earlier: 
\begin{definition}[Conditional Stopping Probability]
The conditional stopping probability  of an audit $\mathcal{A}$ in round $j$ is 
$$\chi_j (\mathcal{A})=\Pr[\mathcal{A}(X)=Correct ~in~round~j~\mid H_a \land \mathcal{A}(X) \neq Correct ~previously]$$
\end{definition}
Experimentally, using our simulations, $\chi_j$ would be estimated by the ratio of the audits that stop in round $j$ to those that ``entered'' round $j$, i.e. those that did not stop before round $j$.

We simulated audits using margins from the 2020 Presidential election, limiting ourselves to pairwise margins for the two main candidates of $0.05$ or larger. Round sizes increase roughly proportional to the inverse
square of the margin, so 
smaller margins are computationally much more expensive to simulate.
For each of these states, we simulated 
$10,000=10^4$ audits assuming the underlying election was as announced ($H_a$),  
and an additional $10,000=10^4$ audits assuming the underlying election was a tie ($H_0$).

We ran simulations for: (a) $90\%$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced and (b) $25\%$ probability of stopping in the first round which is more favorable to \BRAVO. We ran our simulations for up to five rounds. For \Minerva, we chose two round schedules: one where the additional number of ballots chosen in a round is the same as in the previous round (multiplying factor of $1.0$) and the second where the multiplying factor is $1.5$ (this version is integrated into {\em Arlo}, and the multiplying factor was chosen as it roughly ensures a $90\%$ conditional stopping probability in the second round for a first round stopping probability of $90\%$). For both versions of \BRAVO, we chose a single round schedule: each round size has the same conditional stopping probability as the first one.

### 2021-11-02 MultipleRoundSims/sections/05-results.tex
definitions

%Results
For this paper, we simulated audits for 
the 2020 Presidential election
in all US states where the pairwise margin for the two main candidates was at least $5\%$. 
Round sizes increase roughly proportional to the inverse
square of the margin, so 
smaller margins are computationally much more expensive to simulate.
For each of these states, we simulated 
$10,000=10^4$ audits assuming the underlying election was as announced,  
and an additional $10,000=10^4$ audits assuming the underlying election was a tie.

We ran simulations for a $90\%$ probability of stopping in the first round, enabling election officials to be done in the first round with very high probability if the election is as announced.
We ran our simulations for up to five rounds. We assume that the audit results in a hand count if it does not stop in any of $5$ rounds.

\subsection{End-of-Round \BRAVO}
For the EoR \BRAVO simulations, our software estimated and used for each round
the round size that would achieve a $90\%$ probability of stopping if the true election outcome were as announced. In Figure \ref{fig:eor_bravo_sprob}, we display the proportion of EoR \BRAVO audits that stopped in the $j^{th}$ round
to all audits which had not yet stopped before the $j^{th}$ round, for only the first three rounds of the simulations because very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
make it to the $j^{th}$ round. The proportions give an estimate
of the true probability of an EoR \BRAVO audit stopping in the $j$th round,
given that it has not already stopped in a previous round, when the underlying election is as announced. 
We see that, especially in earlier rounds for which 
the values are more representative of true audit behavior, 
our round size predictions are accurate.
In particular, the average across all margins is just above $.9=90\%$ for
all three rounds.

>>>>

%Results
\subsection{End-of-Round \BRAVO}
For the EoR \BRAVO simulations, our software estimated and used for each round
the round size that would achieve a $90\%$ probability of stopping if the true election outcome were as announced. In Figure \ref{fig:eor_bravo_sprob}, we display the proportion of EoR \BRAVO audits that stopped in the $j^{th}$ round
to all audits which had not yet stopped before the $j^{th}$ round, for only the first three rounds of the simulations because very few audits, $(.1)^{j-1}\cdot(10^4)$ on average, 
make it to the $j^{th}$ round. The proportions give an estimate
of the true probability of an EoR \BRAVO audit stopping in the $j$th round,
given that it has not already stopped in a previous round, when the underlying election is as announced. 
We see that, especially in earlier rounds for which 
the values are more representative of true audit behavior, 
our round size predictions are accurate.
In particular, the average across all margins is just above $.9=90\%$ for
all three rounds.
